### 2024-07-18 manuscript/manuscript.tex

\documentclass[
  doc,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{Statistical Reporting Inconsistencies in Experimental
Linguistics}

\shorttitle{Statistical Inconsistencies}

\usepackage{etoolbox}

\authorsnames{Dara Leonard Jenssen Etemady,Timo B. Roettger}

\affiliation{
{Department of Linguistics \& Scandinavian Studies, University of Oslo}}

\leftheader{Etemady and Roettger}

\abstract{Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed
do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim
ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut
aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit
in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
Excepteur sint occaecat cupidatat non proident, sunt in culpa qui
officia deserunt mollit anim id est laborum.}

\keywords{statistics, statcheck, reproducibility}

\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0000-0000-0001}} 
\par{ }
\par{       }
\par{Correspondence concerning this article should be addressed to Timo
B. Roettger, Email: timo.roettger@gmail.com}
}

\makeatletter
\let\endoldlt\endlongtable
\def\endlongtable{
\hline
\endoldlt
}
\makeatother

\urlstyle{same}

\usepackage{annotate-equations}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{1. Introduction}\label{introduction}

What we know about human language and its cognitive underpinnings is
often informed by experimental data. Researchers test theoretical
predictions with these data using statistical tests. Depending on the
results of such tests, researchers make claims for or against
theoretical assumptions.\\
Since they play such an integral part in the reasoning process of
experimentalists, both the data these tests are based on and the
computational procedure of the tests itself should be transparent in
order for other researchers to critically evaluate them. Moreover things
can go wrong. To err is human. Transparent sharing allows others to
detect and correct human error. In the recent years, the quantitative
sciences have seen repeated calls to become more transparent and
reproducible by sharing data and statistical protocols
(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
\citeproc{ref-roettger2019researcher}{Roettger, 2019}). Despite these
calls, sharing of statistical protocols is still rather rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}). If statistical procedures cannot be critically evaluated,
human errors might be left undetected and thus remain uncorrected in the
publication record. And if undetected errors affect the decision
procedure of the analysis, i.e.~whether a hypothesis is rejected or
accepted, these errors might lead to -at best- overconfident, -at worst-
false theoretical conclusions. The present paper will present evidence
that the published literature in experimental linguistics contains a
concerning amount of such statistical errors, a state of affairs which
warrants more rigorous data sharing practices.

\section{2. Statistal reporting
inconsistencies}\label{statistal-reporting-inconsistencies}

The null-hypothesis significance testing (NHST) framework is, to date,
the most dominant statistical framework that researchers use to test
hypotheses in the language sciences
(\citeproc{ref-sondereggera2024advancements}{Sonderegger \& Sóskuthy,
2024}). NHST tests are commonly reported in specific formats which
usually contain the name of the test (e.g.~F, t, χ2), a test statistic,
the degrees of freedom of that test (if applicable), and the p-value,
representing the probability of observing the data (or more extreme
data) given the null hypothesis (i.e.~given that the test statistic is
zero) (see example 1): ~

~

\begin{equation*}
  \tikzmarknode{node1}{(1) }
  \eqnmarkbox[gray]{node2}{F}
  \eqnmarkbox[orange]{node3}{(1,66)}
  \tikzmarknode{node4}{=}
  \eqnmarkbox[purple]{node5}{3.88}
  \tikzmarknode{node6}{, }
  \eqnmarkbox[blue]{node7}{p < .05}
\end{equation*}

\annotate[yshift=1em]{left}{node2}{name of test} 
\annotate[yshift=-1em]{below,left}{node3}{degrees of freedom} 
\annotate[yshift=1em]{right}{node5}{test statistic} 
\annotate[yshift=-1em]{below,right}{node7}{p-value}

~

Since data and statistical protocol sharing remains rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}), interested readers are left with trusting the authors that
the statistical analysis is run and reported correctly. However, the
three sets of indices in (1) have clearly defined mathematical
relationships and can thus be easily checked for consistency. An F test
with the specified degrees of freedom and a test-statistic of 3.88
should result in a p-value of 0.053 which is larger, not smaller, than
0.05. Possible reasons for this inconsistency are manifold: It could be
a typo of the comparison sign, i.e.~the authors meant to use \texttt{=}
or \texttt{\textgreater{}} rather than \texttt{\textless{}}. Any of the
numbers could contain a typo. Sometimes an error might indicate
erroneous rounding (e.g.~0.057 being rounded down to 0.05). Without
access to data and scripts, it remains unclear to the reader what has
caused this inconsistency. Such inconsistencies can be particularly
concerning if the calculated p-value (here 0.057) and the reported
p-values (here \textgreater0.05) are not on the same side of the alpha
threshold. In NHST, p-values below a conventionalized alpha threshold,
most commonly 0.05, are interpreted as evidence that the data are
sufficiently inconsistent with the null hypothesis (``significant'').
P-values above that threshold are considered consistent with the null
hypothesis and practically lead to rejecting the alternative hypothesis
(``non-significant''). In (1) above, the reported p-value suggest a
significant result, the p-value derived from the degrees of freedom and
the test statistic suggest a non-significant result. In the following we
refer to these inconsistencies as ``decision inconsistencies''.

The consistency of these values can be automatically assessed if
statistical tests are reported in an unambiguous format. Recently, a
series of studies used such automatic assessments to evaluate the
prevalence of inconsistent statistical reporting in psychology
(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011},
\citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-claesen2023data}{Claesen et al., 2023};
\citeproc{ref-green2018statcheck}{Green et al., 2018};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016};
\citeproc{ref-nuijten2020statcheck}{Nuijten \& Polanin, 2020};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}), medical
sciences (\citeproc{ref-garcia2004incongruence}{Garcı́a-Berthou \&
Alcaraz, 2004}; \citeproc{ref-van2023comparing}{Van Aert et al., 2023}),
psychiatry (\citeproc{ref-berle2007inconsistencies}{Berle \& Starcevic,
2007}), cyber security studies (\citeproc{ref-gross2021fidelity}{Groß,
2021}), technological education research
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023}), and
experimental philosophy (\citeproc{ref-colombo2018statistical}{Colombo
et al., 2018}). For example, looking at over 250 thousand p-values
published in major psychology journals, Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}) found that around 50\% of
the articles with statistical results contained at least one
inconsistency and around 13\% contained at least one ``decision
inconsistency''. Other studies report on inconsistency rates between 4\%
and 14\%, with between 10\% and 63\% of articles containing at least one
inconsistency and between 3\% and 21\% decision consistencies.

To assess the prevalence of statistical-reporting inconsistencies in
experimental linguistics, the present paper conceptually replicates
Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
p-values reported in eight experimental linguistic journals published
between 2000 and 2023. We explore whether the rate of inconsistency
differs across journals, whether that rate has changed over the course
of the last 20 years and whether there is evidence for bias in these
statistical-reporting inconsistencies. We discuss the results and offer
concrete recommendations for authors, reviewers, and editors to tackle
this problem.

\section{3. Method}\label{method}

All quantitative analyses were conducted using R Core Team
(\citeproc{ref-rmanual}{2021}) and the r packages {[}LIST ALL PACKAGES
from 01\_Analysis.R and add to references{]}

\subsection{3.1. Statcheck}\label{statcheck}

We used the R package wtatcheck (Version 1.4.1-beta.2,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using ``regular expressions''. That way, statcheck can
detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation
tests, and Q tests as long as the test result fulfills three conditions:
(a) the test result is reported completely including the test statistic,
the degrees of freedom (if applicable), and the p-value; (b) the test
result is in the body of the text, i.e.~Statcheck usually misses
reported in tables; and (c) the test result is reported in American
Psychological Association style (\citeproc{ref-APA2020}{American
Psychological Association, 2020}). Given these constraints, Statcheck is
estimated to detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-values and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

\subsection{3.2. Sample}\label{sample}

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the term ``experiment* in title, abstract and/or
keywords. Out of these 100 journals, we selected all journals with at
least 10\% of articles containing the search string''experiment*``. Out
of the remaining 37 journals, we selected those journals that urged APA
formatting either in the main body of the text or specifically regarding
statistics in the author guidelines, resulting in nine remaining
journals. Moreover, to access the articles in .pdf format, the articles
had to be either accessible to us through our library license, or open
access, resulting in a final list of eight journals: Applied
Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
Language Learning and Techology (LLT), Journal of Language and Social
Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
Language Acquisition (SLA).

We included only original research articles within the publication years
of 2000-2023, excluding any book reviews, response articles,
commentaries, editorials, corrigenda, errata, advertisements, etc.
Articles from LAB spanned 2011-2023, while the other journals spanned
2000-2023. Statcheck could not parse 157 articles, likely related to
issues with rendering the Chi-Squared symbol being erroneously converted
from .pdf to .txt. This procedure resulted in 5804 research articles
which were submitted to analysis.

\subsection{3.3. Data availability
statement}\label{data-availability-statement}

All derived data and corresponding R scripts are publicly available
here: LINK.

\section{4. Results}\label{results}

\subsection{4.1 Prevalence of
inconsistencies}\label{prevalence-of-inconsistencies}

The results are summarized in Table~\ref{tbl-summary}. Out of 5804
articles, 3059 articles contained statistical tests that statcheck could
assess (53\%), amounting to 39532 assessable p-values. 5166 p-values
were flagged as inconsistent (13.1\%) out of which 655 were considered
decision inconsistencies (1.7\%) (see Table~\ref{tbl-summary}).

\begin{table}

{\caption{{Number of eligible articles, assessable articles and results,
inconsistencies and decision inconsistencies across all journals.
(Applied Psycholinguistics (APS), Bilingualism: Language and Cognition
(BLC), Linguistic Approaches to Bilingualism (LAB), Language and Speech
(LaS), Language Learning and Techology (LLT), Journal of Language and
Social Psychology (LSP), Journal of Child Language (JCL), and Studies in
Second Language Acquisition (SLA))}{\label{tbl-summary}}}}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0755}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1698}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1887}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1792}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1509}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2358}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Journal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
eligible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessible results
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
inconsistencies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
decision inconsistencies
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
APS & 953 & 690 & 9570 & 1368 & 170 \\
BLC & 964 & 610 & 9093 & 1161 & 120 \\
JCL & 1109 & 529 & 6240 & 750 & 69 \\
LAB & 471 & 133 & 1719 & 234 & 25 \\
LLT & 421 & 111 & 919 & 201 & 61 \\
LSP & 695 & 376 & 4320 & 429 & 60 \\
LaS & 598 & 363 & 4717 & 552 & 86 \\
SLA & 593 & 247 & 2954 & 471 & 64 \\
Total & 5804 & 3059 & 39532 & 5166 & 655 \\
\end{longtable}

\end{table}

The proportion of inconsistencies ranged from 10 to 22\% across journals
(1 to 7\% for decisions inconsistencies) (see
Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
year of publication (see Figure~\ref{fig-year-prop}). On average, 50\%
of assessable articles contained one or more inconsistencies (journals
range from 43 to 56\%) and 13\% contained one or more decisions
inconsistencies (journals range from 11 to 15\%).

\begin{figure}[H]

\caption{\label{fig-stacked-bar}Proportion of articles containing at
least one inconsistency / decision inconsistency. (Applied
Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
Language Learning and Techology (LLT), Journal of Language and Social
Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
Language Acquisition (SLA))}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure1.png}

}

\end{figure}%

\begin{figure}

\caption{\label{fig-year-prop}Proportion of inconsistencies / decision
inconsistencies plotted from 2000 to 2023. Left: rates pooled across
journals; Right: rates per journal. (Applied Psycholinguistics (APS),
Bilingualism: Language and Cognition (BLC), Linguistic Approaches to
Bilingualism (LAB), Language and Speech (LaS), Language Learning and
Techology (LLT), Journal of Language and Social Psychology (LSP),
Journal of Child Language (JCL), and Studies in Second Language
Acquisition (SLA))}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure2.png}

}

\end{figure}%

When examining reported against recalculated p-values for all
inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
spatial patterns (only inconsistencies that are based on `=' comparisons
were included to make them easier to interpret, n = 3779 out of 5166).
First, the center of density of points is located in the bottom left
corner with more p-values reported closer to the standard alpha level of
0.05. This is not surprising, since the quantitative sciences exhibit a
documented publication bias, with hypotheses being more often confirmed
(and thus p \textless{} .05) than not
(\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
is a linear density of points along the diagonal axis which corresponds
to numerically small inconsistencies, some of which might be related to
simple rounding errors. However, comparing the diagonal to the black
line, which represents a linear model predicting recalculated by
reported p-values, we can see a clear divergence of what is expected if
inconsistencies were equally likely in both directions
(i.e.~recalculated p-values were as likely to be smaller than the
reported p-value as larger). The slope of the regression line is flatter
than the diagonal axis which means that, on average, inconsistencies
have a tendency to exhibit smaller reported p-values than their
recalculated counterparts. Similarly, inconsistencies that report the
p-value as being larger or smaller than a reference value (e.g.~p
\textgreater{} 0.05 or p \textless{} 0.05 ) are not equally prevalent.
There were 4.4\% of inconsistencies with p being reported as larger than
a reference but 7.4\% inconsistencies with p being reported as smaller
than a reference (e.g.~p \textless{} 0.05). So even if we assumed these
inconsistencies were merely typos of the comparison sign,
inconsistencies that erroneously report the p-value to be smaller than a
reference value are more frequent than inconsistencies that erroneously
report the p-value to be larger than a reference value. These biases are
also reflected in decisions inconsistencies. Of all decisions
inconsistencies (n = 655), 71\% represent cases in which a reported
significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05).

\begin{figure}

\caption{\label{fig-scatter}Reported vs.~recalculated p-values for
inconsistencies that are based on `=' comparions. Black line
descriptively indicates the linear relationship, indicating a bias
towards lower reported p-values.}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{../plots/figure3.png}

}

\end{figure}%

\section{5. Discussion and
Recommendations}\label{discussion-and-recommendations}

\subsection{5.1. Statistical reporting inconsistencies are
prevalent}\label{statistical-reporting-inconsistencies-are-prevalent}

The present study found a large amount of statistical reporting
inconsistencies across a sample of 5804 experimental linguistic
articles, containing 3059 assessable p-values. 13.1\% of all p-values
were flagged as inconsistent and 1.7\% were flagged as decision
inconsistencies, i.e.~the reported p-value is on the opposite side of
the alpha threshold than the recalculated p-value. On average, 50\% of
assessable articles contained at least one inconsistency and 13\%
contained at least one decision inconsistency.

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent such assessment using Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Previous
studies report on inconsistency rates between 4\% and 14\%, with between
10\% and 63\% of articles containing at least one inconsistency and
between 3\% and 21\% contain at least one decision consistencies.

Even if the prevalence of these inconsistencies could be largely
attributed to inconsequential typos or rounding errors (an assumption we
cannot test without access to the data), the sheer amount of these
inconsistencies that have made it through peer-review should concern us.
They are human errors. If such a substantial amount of errors is found
in plain site, the question naturally arises as to how many errors
during the data analysis itself remain undetected. If the tip of the
iceberg above water is already so large, how large is the body of the
iceberg underneath the surface?

The present examination assessment did not indicate any noticeable trend
over the last 23 years of publications nor did the data suggest large
differences between journals.

Observed inconsistencies were characterized by reported p-values being
on average lower than their recalculated counterparts and the prevalence
of decision inconsistencies was higher for p-values reported as
significant than for those reported as non-significant. These patterns
could indicate a systematic bias in favor of lower p-values in general
and a bias towards significant results in particular. Our data do not
speak to the causes of these biases, but possible reasons include the
following: Researchers might intentionally round down p-values because
they think a lower p-value is more convincing to reviewers and/or
readers. This practice has been admitted to by 1 in 5 surveyed
psychological researchers (\citeproc{ref-john2012measuring}{John et al.,
2012}). Given that a non-trivial amount of quantitative linguists have
admitted to commit questionable research practices (and even fraud)
(\citeproc{ref-isbell2022misconduct}{Isbell et al., 2022}), we cannot
exclude the possibility that some of the inconsistencies were
intentional. It is our strong belief, however, that the majority of
inconsistencies are unintentional.

Researchers might scrutinize non-significant results more than
significant results or are less likely to double check significant
results than non-significant results because results that confirm their
hypothesis feed into their confirmation bias
(\citeproc{ref-nickerson1998confirmation}{Nickerson, 1998}). For
example, Fugelsang et al. (\citeproc{ref-fugelsang2004theory}{2004}) let
researchers evaluate data that are either consistent or inconsistent
with their prior expectations. They showed that when researchers
encounter results that disconfirm their expectations, they are likely to
blame the methodology while results that confirmed their expectations
were rarely critically scrutinized. Alternatively, the observed bias
might be a reflection of publication bias
(\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}) with
(erroneously) reported significant p-values being more likely to be
published than non-significant ones.

\subsection{5.2. Limitations of our
study}\label{limitations-of-our-study}

While we believe our work offers an important contribution to improving
statistical reporting practices in experimental linguistics, there are,
of course, a number of limitations. The present assessment and the
conclusions we can draw from them are limited. First, our sample is
limited to only a subset of experimental linguistic journals. Our sample
is based on a crude criterion of what constitutes an experimental
linguistic journal (see \citeproc{ref-kobrock2023assessing}{Kobrock \&
Roettger, 2023}) and we restricted ourselves further to (for us)
accessible journals which explicitly require APA statistical formatting
in the author guidelines. Thus it is possible that a different selection
of journals would have resulted in different results. However, given
that the inconsistency rates of our study are comparable to similar
studies from other disciplines and that those rates are relatively
stable across journals and time, our findings should be considered
relevant for experimental linguistics at large.

Second, given the constraints on automatically detecting test
statistics, Statcheck misses reported values that either diverge from
APA reporting standards or are reported in tables. However,
inconsistency rates have been shown to be similar for results in APA
format vs.~results that diverge from APA formatting
(\citeproc{ref-bakker2011mis}{Bakker \& Wicherts, 2011};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}).

Third, statcheck slightly overestimates inconsistency rates, because it
might not accurately detect corrections for multiple comparisons
(\citeproc{ref-schmidt2017statcheck}{Schmidt, 2017}). Nuijten et al.
(\citeproc{ref-nuijten2017validity}{2017}), however, show that not only
were there only a small proportion of flagged inconsistencies related to
multiple comparisons, but also that these multiple comparisons
themselves were often erroneously reported. They conclude that
``{[}a{]}ny reporting inconsistencies associated with these tests and
corrections could not explain the high prevalence of reporting
inconsistencies'' (\citeproc{ref-nuijten2017validity}{Nuijten et al.,
2017, p. 27}).

More elaborate automatic tools for the extraction of statistical
information might allow for a more detailed and more accurate assessment
of statistical reporting in the future (e.g.
\citeproc{ref-kalmbach2023rule}{Kalmbach et al., 2023}). Despite its
limitations, Statcheck provides a rough proxy of true inconsistency rate
in the published literature and we hope the reader agrees that the
prevalence of inconsistencies is a state of affairs that should be
reflected upon.

\subsection{5.3. Recommendations for the
field}\label{recommendations-for-the-field}

There are concrete actionable steps the field of experimental
linguistics can make to reduce statistical reporting inconsistencies. In
order to avoid simple copy-and-paste errors related to working in two
separate programs for writing the manuscript and conducting the
statistical analysis, authors should consider `literate programming',
i.e.~an integration of analysis code and prose into a single, dynamic
document (\citeproc{ref-casillas2023opening}{Casillas et al., 2023};
\citeproc{ref-knuth1984literate}{Knuth, 1984}). Several implementations
of literate programming are freely available to researchers including
common R markdown files (Rmd) and Quarto markdown files (qmd). Literate
programming can ensure that values derived from the statistical analysis
are automatically integrated into the manuscript document, avoiding
errors that might happen during a manual transfer from one program to
the other.

Authors should also consider sharing their derived data (i.e.~the
anonymized data table that was analyzed) as well as a detailed
description of their statistical protocol, ideally in form of
reproducible scripts. Sharing reproducible analyses with reviewers
allows the reviewers the reproduce the authors' analyses, possibly
detect errors or even inappropriate statistical choices before
publication, thus improving the quality and robustness of the final
product. Moreover, publicly sharing their analyses has numerous benefits
to the authors themselves beyond error detection: Open data and
materials can facilitate collaboration
(\citeproc{ref-boland2017ten}{Boland et al., 2017}), increase efficiency
and sustainability (\citeproc{ref-lowndes2017our}{Lowndes et al.,
2017}), and are cited more often
(\citeproc{ref-colavizza2020citation}{Colavizza et al., 2020}).

If authors do not share data and scripts, reviewers can check at least
the statistical reporting consistency in the manuscript by using tools
such as statcheck (\citeproc{ref-nuijten2023statcheck}{Nuijten \&
Epskamp, 2023}, http://statcheck.io) or p-checker
(\citeproc{ref-schonbrodt2015p}{Schönbrodt, 2015},
http://shinyapps.org/apps/p-checker/). Reviewers could consider
requesting data and scripts during peer review. Such requests might be
particularly justified when inconsistencies are apparent. Explicitly
requesting to share data might already instill additional care and
quality checks when authors prepare their materials, but also allows the
reviewers to carefully reproduce the results, and critically evaluate
all choices made in the statistical analysis
(\citeproc{ref-sakaluk2014analytic}{Sakaluk et al., 2014}). Recent
evidence suggest experimental linguistics are still characterized by a
pluralism of statistical approaches, even when trying to answer the same
research question (\citeproc{ref-coretta_multidimensional_2023}{Coretta
et al., 2023}). Some of these approaches might be more appropriate than
others (\citeproc{ref-sondereggera2024advancements}{Sonderegger \&
Sóskuthy, 2024}; \citeproc{ref-vasishth2023some}{Vasishth, 2023}), so
more thorough evaluations of how researchers arrive at their statistical
conclusions might elevate their analytical robustness.

Journal editors could explicitly recommend consistency checks with
algorhythms such as statcheck during peer review, a practice that has
been taken up on by several journals from neighboring disciplines
{[}Psychological Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
  accessed on July 15, 2024.}, Advances in Methods and Practices in
Psychological Science\footnote{https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines;
  accessed on on July 15, 2024.}, Stress \& Health Barber
(\citeproc{ref-barber2017meticulous}{2017}){]}. Editors could also
demand, recommend or at least encourage data sharing for publication in
their journal. Data sharing policies have been shown to substantially
increase the reproducibility of analyses
(\citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018}; e.g.
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022}).

Researchers make errors. Researchers have biases. This is who we are as
humans and there is not much we can do about our nature. Being aware of
this fact and how it might affect research might help us to make
possibly negative consequences at worst detectable and best preventable.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-APA2020}
American Psychological Association. (2020). \emph{Publication manual of
the {American Psychological Association}} (7th ed.). Author.
\url{https://doi.org/10.1037/0000173-000}

\bibitem[\citeproctext]{ref-arvan2022reproducibility}
Arvan, M., Pina, L., \& Parde, N. (2022). Reproducibility in
computational linguistics: Is source code enough? \emph{Proceedings of
the 2022 Conference on Empirical Methods in Natural Language
Processing}, 2350--2361.

\bibitem[\citeproctext]{ref-bakker2011mis}
Bakker, M., \& Wicherts, J. M. (2011). The (mis) reporting of
statistical results in psychology journals. \emph{Behavior Research
Methods}, \emph{43}, 666--678.

\bibitem[\citeproctext]{ref-bakker2014outlier}
Bakker, M., \& Wicherts, J. M. (2014). Outlier removal and the relation
with reporting errors and quality of psychological research. \emph{PloS
One}, \emph{9}(7), e103360.

\bibitem[\citeproctext]{ref-barber2017meticulous}
Barber, L. K. (2017). Meticulous manuscripts, messy results: Working
together for robust science reporting. \emph{Stress \& Health},
\emph{33}(2), 89--91.

\bibitem[\citeproctext]{ref-berle2007inconsistencies}
Berle, D., \& Starcevic, V. (2007). Inconsistencies between reported
test statistics and p-values in two psychiatry journals.
\emph{International Journal of Methods in Psychiatric Research},
\emph{16}(4), 202--207.

\bibitem[\citeproctext]{ref-bochynska2023reproducible}
Bochynska, A., Keeble, L., Halfacre, C., Casillas, J. V., Champagne,
I.-A., Chen, K., Röthlisberger, M., Buchanan, E. M., \& Roettger, T.
(2023). Reproducible research practices and transparency across
linguistics. \emph{Glossa Psycholinguistics}, \emph{2}(1).

\bibitem[\citeproctext]{ref-boland2017ten}
Boland, M. R., Karczewski, K. J., \& Tatonetti, N. P. (2017). Ten simple
rules to enable multi-site collaborations through data sharing. In
\emph{PLoS computational biology} (1; Vol. 13, p. e1005278). Public
Library of Science San Francisco, CA USA.

\bibitem[\citeproctext]{ref-buckley2023estimating}
Buckley, J., Hyland, T., \& Seery, N. (2023). Estimating the
replicability of technology education research. \emph{International
Journal of Technology and Design Education}, \emph{33}(4), 1243--1264.

\bibitem[\citeproctext]{ref-caperos2013consistency}
Caperos, J. M., \& Pardo, A. (2013). Consistency errors in p-values
reported in spanish psychology journals. \emph{Psicothema},
\emph{25}(3), 408--414.

\bibitem[\citeproctext]{ref-casillas2023opening}
Casillas, J. V., Constantin-Dureci, G., Rascón, I. A., Shao, J.,
Rodrı́guez, S. A., Gadamsetty, A., Minetti, A., Laungani, K., Thatcher,
J., Gardere, R.-T., et al. (2023). \emph{Opening open science to all:
Demystifying reproducibility and transparency practices in linguistic
research}.

\bibitem[\citeproctext]{ref-claesen2023data}
Claesen, A., Vanpaemel, W., Maerten, A.-S., Verliefde, T., Tuerlinckx,
F., \& Heyman, T. (2023). Data sharing upon request and statistical
consistency errors in psychology: A replication of wicherts, bakker and
molenaar (2011). \emph{Plos One}, \emph{18}(4), e0284243.

\bibitem[\citeproctext]{ref-colavizza2020citation}
Colavizza, G., Hrynaszkiewicz, I., Staden, I., Whitaker, K., \&
McGillivray, B. (2020). The citation advantage of linking publications
to research data. \emph{PloS One}, \emph{15}(4), e0230416.

\bibitem[\citeproctext]{ref-colombo2018statistical}
Colombo, M., Duev, G., Nuijten, M. B., \& Sprenger, J. (2018).
Statistical reporting inconsistencies in experimental philosophy.
\emph{PloS One}, \emph{13}(4), e0194360.

\bibitem[\citeproctext]{ref-coretta_multidimensional_2023}
Coretta, S., Casillas, J. V., Roessig, S., Franke, M., Ahn, B.,
Al-Hoorie, A. H., Al-Tamimi, J., Alotaibi, N. E., AlShakhori, M. K.,
Altmiller, R. M., Arantes, P., Athanasopoulou, A., Baese-Berk, M. M.,
Bailey, G., Sangma, C. B. A., Beier, E. J., Benavides, G. M., Benker,
N., BensonMeyer, E. P., \ldots{} Roettger, T. B. (2023).
Multidimensional {Signals} and {Analytic} {Flexibility}: {Estimating}
{Degrees} of {Freedom} in {Human}-{Speech} {Analyses}. \emph{Advances in
Methods and Practices in Psychological Science}, \emph{6}(3),
25152459231162567. \url{https://doi.org/10.1177/25152459231162567}

\bibitem[\citeproctext]{ref-franco2014publication}
Franco, A., Malhotra, N., \& Simonovits, G. (2014). Publication bias in
the social sciences: Unlocking the file drawer. \emph{Science},
\emph{345}(6203), 1502--1505.

\bibitem[\citeproctext]{ref-fugelsang2004theory}
Fugelsang, J. A., Stein, C. B., Green, A. E., \& Dunbar, K. N. (2004).
Theory and data interactions of the scientific mind: Evidence from the
molecular and the cognitive laboratory. \emph{Canadian Journal of
Experimental Psychology/Revue Canadienne de Psychologie
Exp{é}rimentale}, \emph{58}(2), 86.

\bibitem[\citeproctext]{ref-garcia2004incongruence}
Garcı́a-Berthou, E., \& Alcaraz, C. (2004). Incongruence between test
statistics and p values in medical papers. \emph{BMC Medical Research
Methodology}, \emph{4}, 1--5.

\bibitem[\citeproctext]{ref-green2018statcheck}
Green, C. D., Abbas, S., Belliveau, A., Beribisky, N., Davidson, I. J.,
DiGiovanni, J., Heidari, C., Martin, S. M., Oosenbrug, E., \&
Wainewright, L. M. (2018). Statcheck in canada: What proportion of CPA
journal articles contain errors in the reporting of p-values?
\emph{Canadian Psychology/Psychologie Canadienne}, \emph{59}(3), 203.

\bibitem[\citeproctext]{ref-gross2021fidelity}
Groß, T. (2021). Fidelity of statistical reporting in 10 years of cyber
security user studies. \emph{Socio-Technical Aspects in Security and
Trust: 9th International Workshop, STAST 2019, Luxembourg City,
Luxembourg, September 26, 2019, Revised Selected Papers 9}, 3--26.

\bibitem[\citeproctext]{ref-hardwicke2018data}
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G.
C., Kidwell, M. C., Hofelich Mohr, A., Clayton, E., Yoon, E. J., Henry
Tessler, M., et al. (2018). Data availability, reusability, and analytic
reproducibility: Evaluating the impact of a mandatory open data policy
at the journal cognition. \emph{Royal Society Open Science},
\emph{5}(8), 180448.

\bibitem[\citeproctext]{ref-isbell2022misconduct}
Isbell, D. R., Brown, D., Chen, M., Derrick, D. J., Ghanem, R., Arvizu,
M. N. G., Schnur, E., Zhang, M., \& Plonsky, L. (2022). Misconduct and
questionable research practices: The ethics of quantitative data
handling and reporting in applied linguistics. \emph{The Modern Language
Journal}, \emph{106}(1), 172--195.

\bibitem[\citeproctext]{ref-john2012measuring}
John, L. K., Loewenstein, G., \& Prelec, D. (2012). Measuring the
prevalence of questionable research practices with incentives for truth
telling. \emph{Psychological Science}, \emph{23}(5), 524--532.

\bibitem[\citeproctext]{ref-kalmbach2023rule}
Kalmbach, T., Hoffmann, M., Lell, N., \& Scherp, A. (2023). On the
rule-based extraction of statistics reported in scientific papers.
\emph{International Conference on Applications of Natural Language to
Information Systems}, 326--338.

\bibitem[\citeproctext]{ref-knuth1984literate}
Knuth, D. E. (1984). Literate programming. \emph{The Computer Journal},
\emph{27}(2), 97--111.

\bibitem[\citeproctext]{ref-kobrock2023assessing}
Kobrock, K., \& Roettger, T. (2023). Assessing the replication landscape
in experimental linguistics. \emph{Glossa Psycholinguistics},
\emph{2}(1), 1--28.

\bibitem[\citeproctext]{ref-laurinavichyute2022share}
Laurinavichyute, A., Yadav, H., \& Vasishth, S. (2022). Share the code,
not just the data: A case study of the reproducibility of articles
published in the journal of memory and language under the open data
policy. \emph{Journal of Memory and Language}, \emph{125}, 104332.

\bibitem[\citeproctext]{ref-lowndes2017our}
Lowndes, J. S. S., Best, B. D., Scarborough, C., Afflerbach, J. C.,
Frazier, M. R., O'Hara, C. C., Jiang, N., \& Halpern, B. S. (2017). Our
path to better science in less time using open data science tools.
\emph{Nature Ecology \& Evolution}, \emph{1}(6), 0160.

\bibitem[\citeproctext]{ref-nickerson1998confirmation}
Nickerson, R. S. (1998). Confirmation bias: A ubiquitous phenomenon in
many guises. \emph{Review of General Psychology}, \emph{2}(2), 175--220.

\bibitem[\citeproctext]{ref-nuijten2017validity}
Nuijten, M. B., Assen, M. A. van, Hartgerink, C., Epskamp, S., \&
Wicherts, J. M. (2017). \emph{The validity of the tool {``statcheck''}
in discovering statistical reporting inconsistencies}.

\bibitem[\citeproctext]{ref-nuijten2023statcheck}
Nuijten, M. B., \& Epskamp, S. (2023). \emph{Statcheck: Extract
statistics from articles and recompute p-values(1.4. 1-beta. 2){[}r{]}}.

\bibitem[\citeproctext]{ref-nuijten2016prevalence}
Nuijten, M. B., Hartgerink, C. H., Van Assen, M. A., Epskamp, S., \&
Wicherts, J. M. (2016). The prevalence of statistical reporting errors
in psychology (1985--2013). \emph{Behavior Research Methods}, \emph{48},
1205--1226.

\bibitem[\citeproctext]{ref-nuijten2020statcheck}
Nuijten, M. B., \& Polanin, J. R. (2020). {``Statcheck''}: Automatically
detect statistical reporting inconsistencies to increase reproducibility
of meta-analyses. \emph{Research Synthesis Methods}, \emph{11}(5),
574--579.

\bibitem[\citeproctext]{ref-rmanual}
R Core Team. (2021). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-roettger2019researcher}
Roettger, T. B. (2019). Researcher degrees of freedom in phonetic
research. \emph{Laboratory Phonology}, \emph{10}(1).

\bibitem[\citeproctext]{ref-sakaluk2014analytic}
Sakaluk, J., Williams, A., \& Biernat, M. (2014). Analytic review as a
solution to the misreporting of statistical results in psychological
science. \emph{Perspectives on Psychological Science}, \emph{9}(6),
652--660.

\bibitem[\citeproctext]{ref-schmidt2017statcheck}
Schmidt, T. (2017). \emph{Statcheck does not work: All the numbers.
Reply to nuijten et al.(2017)}.

\bibitem[\citeproctext]{ref-schonbrodt2015p}
Schönbrodt, F. D. (2015). \emph{P-checker: One-for-all p-value
analyzer}. \url{http://shinyapps.org/apps/p-checker/}.

\bibitem[\citeproctext]{ref-sondereggera2024advancements}
Sonderegger, M., \& Sóskuthy, M. (2024). \emph{Advancements of phonetics
in the 21st century: Quantitative data analysis}.

\bibitem[\citeproctext]{ref-sterling1959publication}
Sterling, T. D. (1959). Publication decisions and their possible effects
on inferences drawn from tests of significance---or vice versa.
\emph{Journal of the American Statistical Association}, \emph{54}(285),
30--34.

\bibitem[\citeproctext]{ref-van2023comparing}
Van Aert, R. C., Nuijten, M. B., Olsson-Collentine, A., Stoevenbelt, A.
H., Van Den Akker, O. R., Klein, R. A., \& Wicherts, J. M. (2023).
Comparing the prevalence of statistical reporting inconsistencies in
COVID-19 preprints and matched controls: A registered report.
\emph{Royal Society Open Science}, \emph{10}(8), 202326.

\bibitem[\citeproctext]{ref-vasishth2023some}
Vasishth, S. (2023). Some right ways to analyze (psycho) linguistic
data. \emph{Annual Review of Linguistics}, \emph{9}(1), 273--291.

\bibitem[\citeproctext]{ref-veldkamp2014statistical}
Veldkamp, C. L., Nuijten, M. B., Dominguez-Alvarez, L., Van Assen, M.
A., \& Wicherts, J. M. (2014). Statistical reporting errors and
collaboration on statistical analyses in psychological science.
\emph{PloS One}, \emph{9}(12), e114876.

\bibitem[\citeproctext]{ref-wicherts2011willingness}
Wicherts, J. M., Bakker, M., \& Molenaar, D. (2011). Willingness to
share research data is related to the strength of the evidence and the
quality of reporting of statistical results. \emph{PloS One},
\emph{6}(11), e26828.

\end{CSLReferences}

\appendix

\section{Title for Appendix}\label{title-for-appendix}

\end{document}